Practical Optional Types for Clojure

### 2015-02-28 overview.tex
3.4
cites

We chose to build on the ideas and implementation
of Typed Racket to implement a type system targeting Clojure for several reasons.
Initially, the similarities between Racket and Clojure drew us to
investigate the effectiveness of repurposing occurrence typing
for a Clojure type system---both languages share a Lisp heritage,
similar standard functions 
(for instance \clj{map}
in both languages is variable-arity)
and idioms.
While Typed Racket is gradually typed and has sophisticated
dynamic semantics for cross-language interaction, we 
chose to first implement
the static semantics
with the hope to extend Typed Clojure to be gradually typed at a future date.
Finally,
Typed Racket's combination of bidirectional checking
and occurrence typing presents a successful model for 
type checking dynamically typed programs without compromising
soundness, which is appealing over success typing~\cite{todo-dialyzer-paper}
which cannot prove strong properties about programs
and soft typing~\cite{successtyping-todo}
which has proved too complicated in practice.

@@

\begin{figure}
\begin{exmp}
\inputminted[firstline=6,lastline=22]{clojure}{code/demo/src/demo/eg7.clj}
\label{example:multidispatch}
\end{exmp}
\end{figure}

@@

\begin{figure}
\begin{exmp}
\inputminted[firstline=6,lastline=20]{clojure}{code/demo/src/demo/eg8.clj}
\label{example:final}
\end{exmp}
\end{figure}

>>>>

We chose to build on the ideas and implementation
of Typed Racket to implement a type system targeting Clojure for several reasons.
Initially, the similarities between Racket and Clojure drew us to
investigate the effectiveness of repurposing occurrence typing
for a Clojure type system---both languages share a Lisp heritage,
similar standard functions 
(for instance \clj{map}
in both languages is variable-arity)
and idioms.
While Typed Racket is gradually typed and has sophisticated
dynamic semantics for cross-language interaction, we 
chose to first implement
the static semantics
with the hope to extend Typed Clojure to be gradually typed at a future date.
Finally,
Typed Racket's combination of bidirectional checking
and occurrence typing presents a successful model for 
type checking dynamically typed programs without compromising
soundness, which is appealing over success typing~\cite{Lindahl:2006:PTI}
which cannot prove strong properties about programs
and soft typing~\cite{CF91}
which has proved too complicated in practice.

@@

@@

\begin{figure}
\begin{exmp}
\inputminted[firstline=6,lastline=22]{clojure}{code/demo/src/demo/eg7.clj}
\label{example:multidispatch}
\end{exmp}
\begin{exmp}
\inputminted[firstline=6,lastline=20]{clojure}{code/demo/src/demo/eg8.clj}
\label{example:final}
\end{exmp}
\caption{Multimethod Examples}
\end{figure}
